---
title: "Analyzing TimeTracker output from art"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Analyzing TimeTracker output from art}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message = FALSE}
library(dplyr)
library(artsupport)
```

# Overview

The _art_ [`TimeTracker`](https://cdcvs.fnal.gov/redmine/projects/art/wiki/TimeTracker)
service records a variety of timing information in a running _art_ framework program,
and optionally writes that data to an [SQLite](https://sqlite.org) database.

These databases contain three tables:

1. *TimeEvent*, which has one record per event, giving the processing time of the event *except* for writing output;
2. *TimeSource*, which has one one record per event, giving the name and execution time of the source; and
3. *TimeModule*, which has one record per module per event, giving the identity of the module and its execution time on each event.

We'll be looking at a small database produced by running an ICARUS simulation program.

```{r}
dbfile <- here::here("tests/testthat/timing-with-output-module.db")
```

# Event timing data

# Source timing data

Source timing data can be loaded using `load_source_timing`:
```{r}
sources <- load_source_timing(dbfile)
sources
```


It is important to note that `RootInput`, the source used to read _art_/Root files,
is designed to support delayed loading of data products.
This means that `RootInput` only reads a small amount of metadata for each event,
and thus it appears very fast.
The actual reading of a data product is delayed until the first request for that data product is made by a module.
If you want to determine the time taken to read the input, you can set the configuration to do so:
```
source.delayedReadEventProducts: false
```
But note that this may cause a degradation in the oveall processing speed of the program.

# Module timing data

Module timing data can be loaded using `load_module_timing`:

```{r}
modules <- load_module_timing(dbfile)
modules
```

The *source* module is not included in the `TimeModule` table.
If you want to include source in the data read by `load_module_timing`, set the parameter `include_source = TRUE`:

```{r}
all_modules <- load_module_timing(dbfile, include_source = TRUE)
```



The `TimeTracker` records two values on each event for output modules;
they are distinguished by one having the suffix `(write)` in the `ModuleType` column.

```{r}
modules %>% filter(Path == "end_path")
```

The record without the `(write)` suffix contains the time taken by the event-processing function of the output module.
This does *not* include the time actually taken to write to the output.
In includes things like the running of a `ResultsProducer`, which happens in the context of an output module.
The record with the `(write)` suffix is the time taken to write the output.

